Automated Formal Synthesis of Digital Controllers for State-Space Physical Plants

### 2017-01-18 cav2017/cav2017.tex
minor changes

A discrete-time system is asymptotically stable if and only if all 
characteristic polynomial roots (i.e., eigenvalues) are inside of 
the unity circle, i.e., their absolute value are less than 
one~\cite{astrom1997computer}. For a closed-loop system as~\eqref{eq:closedloopss}, 
the characteristic polynomial $P(z)$ can be computed as follows:
\begin{equation}
P(z)= \det( z I_{n} - A_d + B_d K ),
\end{equation}
where $I_{n}$ is the n-th order identity matrix.

Jury's criterion provides necessary and sufficient conditions for the eigenvalues 
to be within the unit circle and, consequently, for the system to be stable.  Given a 
polynomial $P(z)$ of the form
$$
P(z) = a_{0}z^{N} + a_{1}z^{N-1} + ... a_{N-1}z + a_{N} = 0, a_{0}\neq 0,
$$
Jury's criterion explores solely the coefficients ($a_{0},a_{1},...,a_{N-1}$) 
of $P(z)$ for checking the validity of the stability property (represented by 
$\phi_{stability}$).

Jury stability test is already explained in the control system 
literature~\cite{fadali}. This study, however, limits itself to 
explain the encoding of Jury's criterion for the formal synthesis purpose. 
For the stability  test procedure, the following Jury matrix 
$M=[m_{ij}]_{(2N-2)\times N}$ is built from $P(z)$ coefficients:
$$
M=\left[
\begin{matrix}
  V^{(0)} \\
  V^{(1)} \\
  \vdots \\
  V^{(N-2)}
 \end{matrix}
\right]\mbox{,}
$$
\noindent where $V^{(k)}=[v^{(k)}_{ij}]_{2\times N}$ such that
$$
v^{(0)}_{ij}=\begin{cases} 
a_{j-1}, & \mbox{if } i=1 \\   v^{(0)}_{(1)(N-j+1)}, & \mbox{if } i=2 
\end{cases}\mbox{, and} 
$$
$$
v^{(k)}_{ij}=\begin{cases} 
0, & \mbox{if } j>n-k\\
v^{(k-1)}_{1j}-v^{(k-1)}_{2j}\cdot\frac{v^{(k-1)}_{11}}{v^{(k-1)}_{21}}, & \mbox{if } j\leq n-k  \mbox{ and } i=1 \\
v^{(k)}_{(1)(N-j+1)}, & \mbox{if } j\leq n-k \mbox{ and } i=2 \\
\end{cases} \mbox{,}
$$
\noindent where $k\in\mathbb{Z}$, such that $0<k<N-2$. $P(z)$ is the 
characteristic polynomial of a stable system if and only if the following 
four propositions hold:
\begin{itemize}
\item $R_{1}$: $S(1)>0$;
\item $R_{2}$: $(-1)^{N}P(-1)>0$;
\item $R_{3}$: $\vert{a_{0}}\vert <a_{N}$;
\item $R_{4}$: $m_{11}>0\iff m_{31}\wedge m_{51}\wedge \dots \wedge m_{(2N-3)(1)}$.
\end{itemize}

@@

Finally, we also require a safety specification $\Psi(\vec{u},\vec{y})$, which is
given as a set of linear constraints over the inputs and outputs.

>>>>

A discrete-time system is asymptotically stable if and only if all the roots 
of the characteristic polynomial ({\it i.e.}, the eigenvalues) are inside of 
the unity circle, {\it i.e.}, their absolute value are less than 
one~\cite{astrom1997computer}. For a closed-loop system as~\eqref{eq:closedloopss}, 
the characteristic polynomial $P(z)$ should be computed as follows:
\begin{equation}
P(z)= \det( z I_{n} - A_d + B_d K ),
\end{equation}
where $I_{n}$ is the n-th order identity matrix.

Jury's criterion provides necessary and sufficient conditions for the eigenvalues 
to be within the unit circle and, consequently, the system is stable.  Given a 
polynomial $P(z)$ of the form
$$
P(z) = a_{0}z^{N} + a_{1}z^{N-1} + ... a_{N-1}z + a_{N} = 0, a_{0}\neq 0,
$$
Jury's criterion explores solely the coefficients ($a_{0},a_{1},...,a_{N-1}$) 
of $P(z)$ for checking the validity of the stability property $\phi_{stability}$.

In particular, Jury stability test is already explained in the control system 
literature ({\it e.g.}~\cite{fadali}). This study, however, limits itself to 
explain the encoding of Jury's criterion for the formal synthesis purpose. 
For the stability  test procedure, the following Jury matrix 
$M=[m_{ij}]_{(2N-2)\times N}$ is built  from $P(z)$ coefficients:
$$
M=\left[
\begin{matrix}
  V^{(0)} \\
  V^{(1)} \\
  \vdots \\
  V^{(N-2)}
 \end{matrix}
\right]\mbox{,}
$$
\noindent where $V^{(k)}=[v^{(k)}_{ij}]_{2\times N}$ such that
$$
v^{(0)}_{ij}=\begin{cases} 
a_{j-1}, & \mbox{if } i=1 \\   v^{(0)}_{(1)(N-j+1)}, & \mbox{if } i=2 
\end{cases}\mbox{, and} 
$$
$$
v^{(k)}_{ij}=\begin{cases} 
0, & \mbox{if } j>n-k\\
v^{(k-1)}_{1j}-v^{(k-1)}_{2j}\cdot\frac{v^{(k-1)}_{11}}{v^{(k-1)}_{21}}, & \mbox{if } j\leq n-k  \mbox{ and } i=1 \\
v^{(k)}_{(1)(N-j+1)}, & \mbox{if } j\leq n-k \mbox{ and } i=2 \\
\end{cases} \mbox{,}
$$
\noindent where $k\in\mathbb{Z}$, such that $0<k<N-2$. $P(z)$ is the 
characteristic polynomial of a stable system if and only if the following 
four propositions hold:
\begin{itemize}
\item $R_{1}$: $S(1)>0$;
\item $R_{2}$: $(-1)^{N}P(-1)>0$;
\item $R_{3}$: $\vert{a_{0}}\vert <a_{N}$;
\item $R_{4}$: $m_{11}>0\iff m_{31}\wedge m_{51}\wedge \dots \wedge m_{(2N-3)(1)}$.
\end{itemize}

@@

Finally, we also require that the closed-loop systems meets safety specifications. A safety specification establish bounds for the states, such that he controller $K$ must maintain the states within these bounds for any initial state.

The safety property (represented by $\phi_{safety}$) is encoded as follows:
\begin{equation}
\label{eq:safetyliteral}
\phi_{safety}\iff \forall k\geq 0, \bigwedge_{i=1}^{n}{x_{i}^{-} \leq x_{i,k} \leq x_{i}^{+}},
\end{equation}
where $x_{i}^{-}$ and $x_{i}^{+}$ are respectively the lower and upper bound of the $i$-th state $x_{i}$ at the $k$-th instant, i.e., the states will always be within the $n$-dimensional hyperbox defined by the safety specifications (bounds).

The closed-loop state-space system have to be unrolled until a specific depth ($k_{bound}$) to ensure $\phi_{safety}$. The limit $k_{bound}$ is the number of iterations sufficient so that the rate of variation (derivative) of the states is negligible ensuring the boundaries will not be violated for any $k>k_{bound}$. \commentib{The k-bound estimative must be explained.}

For any $k\leq k_{bound}$ the states must be computed considering the FWL effects in all operations related to the control action~\ref{eq:controlaction} and disregarding it in the state equation ~\eqref{eq:plant}. Thus, the states are computed using the following equations:
\begin{equation}
\label{eq:ssunroll}
\left\lbrace
\begin{array}{l}
u_{k}=\mathcal{F}_{\langle I,F \rangle}(r_{k}-\mathcal{F}_{\langle I,F \rangle}(\mathcal{F}_{\langle I,F \rangle}(K)\cdot\mathcal{F}_{\langle I,F \rangle}(x_{k})))\\
x_{k+1}=Ax_{k}+Bu_{k}
\end{array}\right.
\end{equation}

The input sequence $u_{k}$ is non-deterministic with known bounds $u^{+}$ and $u^{-}$, such that ${u^{-} \leq u_{k} \leq u^{+}},\forall k\geq 0$.
